%!TEX root = thesis.tex

%% Write your acknowledgements in this file. If you do not want to acknowledge anyone,
%% you can delete this file and comment out the corresponding part in the "thesis.tex"
%% file.


This dissertation is the result of multiple years of work and the involvement of many people. There are no words to express the gratitude and appreciation to every single person that supported me throughout the last 7 years at NYU. Thanks to all of you, I finish this stage in life with a heart filled with joy and hope for things to come. 

First and foremost, I want to thank my dissertation committee. Pablo Querubin, thank you for believing in me since we first met, for giving me the opportunity to teach side-by-side with you for five years, for always being patient with me, for spending countless hours to talk about my research and even job opportunities, for always doing so with a big smile. Thank you for your advice inside and outside the classroom, for leading me to new research opportunities, for pushing my research, for being my mentor and friend. I will never forget the words you gave to the Latin American Politics class in the last lecture of Spring of 2017: you spoke of how the class revolved around portraying the faults of Latin America but that students should not loose sight of the beauties of the region, and the promise of hope for it. Your love for Latin America is something I will carry on for the rest of my life. Cyrus Samii, there is no way I can express how much I thank you for all your support these years. There were times in the Ph.D. program where I thought I couldn't make it but it was your advice and guidance throughout these years that help me build confidence in my research. I will never forget the days spend in Morocco, your help to meet new faculty members, your methodological teaching, your mentoring and support. Hye Young You, the first time I met you, you had read all of my work in such a thorough way that left me amazed. Your insights and clarity have been extremely helpful, and much of what I could do in this dissertation was thanks to those meetings with you. If the kindness and patience you have showed to me was present in every faculty member, academia would be the best of places in the world. Tara Slough, as I have told you before, every time I see you participate in a seminar, deliver comments to students or in conferences, and when you have spoken directly to my research you have always provided the best of feedback. I have long admired your research and work ethic. Begin honest, I wish I could be like you, and I'm so happy to have you as part of my committee. Juan Vargas, mi mentor y mi amigo, if there is someone to truly thank throughout all these years is you. Since the Summer of 2015 you opened your arms to me as a collegue, and few days after as friend for life. There is no one I have learned from as you, who truly built up my foundations as a researcher, who opened up research opportunities to me that I would have never dreamed about. 

I also want to thank the faculty members at NYU and Princeton University. Neal Beck, there is no man in academia like you, there is no one that loves students as you do. It is becasue of you that a boy in a Masters program with hopes of joining a Ph.D. program achieved his dream as is now about to finish it. I will be eternally grateful for your support and friendship. Jake Shapiro, I met you at the end of my first year at the Masters at NYU. Since then my life changed and I couldn't even imagine everything that happened since, and much of it was due to your mentoring, support, and friendship. You are a dreambuilder, and I have achieved many of my dreams by being under your wing. I'm truly thankful to you and excited for the work to come in the following years. Leonard Wantchekon, thank you for sharing some of your passion for research and how to better the world we are living in. Your trust in me and all the research opportunities you have shared with me cannot be repaid.

Thank you for all the faculty members at NYU that shared their knowledge and advice during my doctoral studies. Thank you Dimitri Landa, Peter Rosendorff, Shanker Satyanath, Nicole Simonelli,  Amanda Kennard, Tiberiu Dragu, Mike Gilligan, David Stasavage, and Julia Payson.  

God has given me the best friends: I am truly the happiest man of the world because of you. Gordos, Gabriel y Gizeh, mis amigos del alma, los tres mosqueteros hasta el final, verdaderamente me han enseñado que el hierro con hierro se afila. Alan, el hermano que nunca tuve y que me ha acompañado hasta este día. Si en alguien dejara mi vida sería en tus manos. Así como tus logros son mis logros, espero \'este sea uno tuyo pues no pude haberlo alcanzado sin tu amistad. Andresiño -my fish!- y Juanda, mis dos hermanos de otros padres, dos regalos que Dios me dió de pura gracia y que serán mi familia por siempre. Nick, I never thought to find a friend like you in the States, and to admire a man the way I admire you. You are one of God's best gifts I have in my life. I have missed you since you left NYC. Prabin, your heart is the biggest of all, and your friendship a true gift. To both of you, Nick and Prabin, I couldn´t have done the Ph.D. without you, you were the greatest support, and the truest of friends. Reed, I wish we could have met earlier in the progam but I thank God for that trip to China, for the endless talks we have had since, for your friendship, your guidance and mentoring. We are brothers for life. To all my friends at NYU, thank you: Will, Antonella, Mateo, Mariana, Dongil, Abraham, Felipe, Martin, Patrick, Arina, Maria, Anne, Giacomo, Athena, Rajeshwari, Lucia, Ye, and Stephanie.

To my friends in NYC, thank you for being present throughout these years, for all your prayers and support. Rich, my dear friend and brother, you have been and advisor and the best of friends. Truly, you have showed me that there is a friend that is closer than a brother. Jing, there is no day that I miss you in NY. God blessed me with your life since we met. To all the members of the Stuy CG and Redeemer, Fei Mei, Sam, Melody, Miguelo, Emilia, Diane, Becks, Audrey, and many, many more. To all of you and even the ones I don't name, thank you for your frienship and your love. You are my NYC family, and there is no better family than you all. 

To my students from the IR Honors, Latin American Politics, and Political Economy of Development classes at NYU, thank you. Back when I was living in Mexico and before coming to the States I had a dream of becoming an advisor to kings and queens. For years I thought I would achieve that by taking a job in public office. You, however, thought me that by teaching I was already fulfilling the biggest dream in my heart. You, my kings and queens, are my joy and glory. 

Finally, and most importantly, I'm grateful to my family. Mamá eres a quien más admiro en el mundo y cuyo amor me sostuvo todos estos años: mi gran consejera y mi gran amiga. Arela y Lala, mis amores, mis hermanitas hermosas que con su amor me apoyaron día tras día, y me hacen querer ser mejor cada día. A las tres sepan que mis sueños cumplidos son suyos, mis logros sus logros, mi honor su honor, pues no hay nada que que hubiera logrado sin ustedes. Su amor es mi motor y el regalo más grande de Dios en esta vida. Esta disertación se las dedico a ustedes. Las amo.

God, I have fought the good fight, I have finished the race, I have kept the faith. You have lifted my head up high, you have sustained me and loved me like nobody has. My life is yours, this degree is yours, and to you, who is able to do immesurably more than all we ask or imagine, to you God be the glory forever and ever. 

\noindent
\makebox[\textwidth]{\hfill\makebox[3in]{\hfill Rafael J. Ch Duran\hfill}}
\makebox[\textwidth]{\hfill\makebox[3in]{\hfill\graddate\hfill}}

