%!TEX root = thesis.tex

I’m a Ph.D. Candidate in the Politics Department at New York University. My research interests lie in Comparative Politics and Methods, and include the study of Governance, Conflict, and State Capture. I have a regional specialization in developing countries, particularly Latin America.

I received NYU’s Dean’s Dissertation Fellowship 2020-2021. My research has also been supported by organizations including the Fulbright-Garcia Robles Program, the Development Bank of Latin America (CAF), the Center for Studies in Security and Drugs (CESED) from Universidad Los Andes, among others. My scholarly work has been published in the American Political Science Review, the Journal of Urban Economics and I have been a referee for the American Political Science Review, Comparative Political Studies, and Latin American Research Review.

While in the Ph.D. program at NYU, I also served as Principal Investigator for the Latin America Development Bank´s (CAF) “Slum Observatory” Project, and worked on CAF’s leading chapter of the flagship annual Economy and Development Report 2017 on Urban Development in Latin America. In 2015 I served as a Research Assistant at the Empirical Studies of Conflict (ESOC) at Princeton University. Prior to NYU, I served as Director of the Economic and Public Security Research Division at the Center of Research for Development, A.C. (CIDAC), and as a Visiting Professor at the Political Science Department of the Instituto Tecnológico Autónomo de México (ITAM) where I thought Political Methodology. I received my BA in Economics from ITAM with the highest distinction. 
